Biomass predictions and VOCC

### 2019-08-09 analysis/VOCC/4-vocc-by-spp.Rmd
beginning to simplify vocc-biomass merge code

Species

@@

#climate <- "temperature"
temp_threshold <- c(0.75)
optimal_temp <- 6.5 # yellow-eye 4

@@

climate <- "do and temperature"
max_thresholds <- c(Inf, 0.75)
min_thresholds <- c(0.25, Inf)
optimal_value <- c(2, 6.5)
```

@@

Run all subsequent code...
```{r global_options, include=FALSE}
library(dplyr)
library(ggplot2)
library(gfplot)
library(gfdata)
library(sdmTMB)
library(gfranges)

knitr::opts_chunk$set(
  fig.width = 11, fig.height = 8.5,
  echo = FALSE, warning = FALSE, message = FALSE
)
```

@@

```{r}
rm(vocc)
try({
  vocc <- readRDS(paste0("data/vocc-", ssid_string,"-", climate_string, "-", optimal_string, "-",
    min_string,"to", max_string, "-", start_time, "-", delta_t_total, ".rds"))
})

@@

vocc_plot <- plot_vocc(vocc,
  vec_aes = "distance",
  #max_vec_plotted = 100,
  min_vec_plotted = input_cell_size,
  NA_label = ".",
  fill_col = "diff",
  fill_label = paste("Change from\n", start_time, "to", start_time + delta_t_total, ""),
  raster_alpha = 1,
  vec_alpha = 0.25,
  axis_lables = FALSE,
  transform_col = no_trans#,
  #raster_limits = c(5, 13)
)
vocc_plot + ggtitle(climate)
}

@@

vocc_plot <- plot_vocc(vocc,
  vec_aes = "distance",
  #max_vec_plotted = 100,
  min_vec_plotted = input_cell_size,
  NA_label = ".",
  fill_col = "diff",
  fill_label = paste("Change from\n", start_time, "to", start_time + delta_t_total, ""),
  raster_alpha = 1,
  high_fill = "Steel Blue 4",
  low_fill = "Red 3",
  vec_alpha = 0.25,
  axis_lables = FALSE,
  transform_col = no_trans#,
  #raster_limits = c(5, 13)
)
vocc_plot + ggtitle(climate)
}

@@

vocc_plot_temp <- plot_vocc(vocc,
  vec_aes = "distance",
  #max_vec_plotted = 100,
  min_vec_plotted = input_cell_size,
  NA_label = ".",
  fill_col = "temp_diff",
  fill_label = paste("Change from\n", start_time, "to", start_time + delta_t_total, ""),
  raster_alpha = 1,
  vec_alpha = 0.25,
  axis_lables = FALSE,
  transform_col = no_trans#,
  #raster_limits = c(5, 13)
)+ ggtitle("temperature")

@@

vocc_plot_do <- plot_vocc(vocc,
  vec_aes = "distance",
  #max_vec_plotted = 100,
  min_vec_plotted = input_cell_size,
  NA_label = ".",
  fill_col = "do_diff",
  fill_label = paste("Change from\n", start_time, "to", start_time + delta_t_total, ""),
  raster_alpha = 1,
  high_fill = "Steel Blue 4",
  low_fill = "Red 3",
  vec_alpha = 0.25,
  axis_lables = FALSE,
  transform_col = no_trans#,
  #raster_limits = c(5, 13)
) + ggtitle("do")

@@

start <- biomass_change %>% 
  rename(start_density = var_1_s, end_density = var_1_e) %>% 
  select(icell, x, y, start_density, end_density)

end <- biomass_change %>% 
  rename(start_density = var_1_s, 
         end_density = var_1_e, 
         target_X = x, target_Y = y) %>% 
  select(icell, start_density, end_density, target_X, target_Y)

source_density <- biomass_change %>% rename(source_density = var_1_s) %>% select(source_density, x, y)

if (length(variable_names) == 1) {
  vectors <- vocc %>%
    mutate(
      start_cell = icell, 
      vector_id = paste0(id, "_to_", tid),
      initial_value = var_1_s, 
      source_end_value = var_1_e, 
      target_value = as.numeric(target_values)
    ) %>%
    filter(distance > 1)  %>%
    select(vector_id, start_cell, x, y, target_X, target_Y, distance, 
      initial_value, source_end_value, target_value, 
      km_per_decade, timespan)

  if (climate == "do") vectors <- vectors %>% filter(source_end_value < (optimal_do - do_threshold))
  if (climate == "temperature") vectors <- vectors %>% filter(source_end_value > (optimal_temp + temp_threshold))

  vectors <- vocc %>%
    mutate(
      start_cell = icell, vector_id = paste0(id, "_to_", tid),
      initial_value = s, source_end_value = e, target_value = target_values
    ) %>%
    filter(distance > 1) %>%
    select(vector_id, start_cell, x, y, target_X, target_Y, distance, 
      initial_value, source_end_value, target_value, 
      km_per_decade, timespan)

  source_end <- purrr::map_df(strsplit(vectors$source_end_value, " "), function(x) data.frame(t(as.numeric(x))))
  vectors <- cbind(vectors, source_end)

  for (j in seq_along(variable_names)){
  browser()
    vectors <-  filter( vectors, paste0("X",j) < (optimal_value[j] - min_thresholds[j]) )
  }

@@

# get mean and sd for densities without cell duplication
icells <- vocc_by_sp[[i]] %>% distinct(icell, .keep_all = TRUE)
mean.change <- mean(icells$perc_change, na.rm = TRUE)
sd.change <- sd(icells$perc_change, na.rm = TRUE)

@@

# calculate standardized density varibles
vocc_by_sp[[i]] <- vocc_by_sp[[i]] %>% 
  mutate(
    perc_change = (end_density - start_density)/start_density, 
    cell_type = as.factor(cell_type),
    std_change = (perc_change)/sd.change, 
    std_source_density = (source_density - mean.density)/(sd.density*2), 
    mean_change = mean.change, sd_change = sd.change,
    mean_source_density = mean.density, sd_density = sd.density,
# set distance intercept at cell size
    distance = (ifelse(distance > 100, 100, distance)) - input_cell_size, 
    distance2 = distance*distance) %>% 
  rename(X = x, Y = y)

@@

vectors <- vectors %>% rename(
  source_end_climateA = X1, source_end_climateB = X2
  )

mean.climateA <- mean(icells$source_end_climateA, na.rm = TRUE)
sd.climateA <- sd(icells$source_end_climateA, na.rm = TRUE)
mean.climateB <- mean(icells$source_end_climateB, na.rm = TRUE)
sd.climateB <- sd(icells$source_end_climateB, na.rm = TRUE)

vocc_by_sp[[i]] <- vocc_by_sp[[i]] %>% 
  mutate(
    mean_climateA = mean.climateA, sd_climate = sd.climateA,
    std_end_climateA = (source_end_climateA - mean.climateA)/(sd.climateA*2),
    std_end_climateA2 = std_end_climateA^2,
    mean_climateB = mean.climateB, sd_climate = sd.climateB,
    std_end_climateB = (source_end_climateB - mean.climateB)/(sd.climateB*2),
    std_end_climateB2 = std_end_climateB^2
    )
} else {
 stop("Doesn't currently support testing vectors bases on more than two climate variables at once.")
}
}

@@

# glimpse(vocc_by_sp[[i]])
# trim spatial outliers 
# Xmax <- quantile(vocc_by_sp[[i]]$X, 0.9995)
# Xmin <- quantile(vocc_by_sp[[i]]$X, 0.005)
# std_max <- quantile(vocc_by_sp[[i]]$std_change, 0.9995)
# 
# vocc_by_sp[[i]] <- vocc_by_sp[[i]] %>% filter ( X >= Xmin & X <= Xmax )

@@

Check distance threshold of 50 looks reasonable
```{r}
data <- vocc_by_sp[["adult"]]
#data <- filter(vocc_by_sp[["adult"]], distance > 1 & distance < 100)

>>>>

Species
```{r global_options, include=FALSE}
# library(dplyr)
# library(ggplot2)
# library(gfplot)
# library(gfdata)
# library(sdmTMB)
# library(gfranges)
# 
# knitr::opts_chunk$set(
#   fig.width = 11, fig.height = 8.5,
#   echo = FALSE, warning = FALSE, message = FALSE
# )
```

@@

climate <- "temperature"
temp_threshold <- c(0.75)
optimal_temp <- 6.5 # yellow-eye 4

@@

#climate <- "do and temperature"
max_thresholds <- c(Inf, 0.75)
min_thresholds <- c(0.25, Inf)
critical_thresholds <- c(-0.25, 0.75)
optimal_value <- c(2, 6.5)
```

@@

Run all subsequent code...

@@

```{r}
rm(vocc)
# try({
#   vocc <- readRDS(paste0("data/vocc-", ssid_string,"-", climate_string, "-", optimal_string, "-",
#     min_string,"to", max_string, "-", start_time, "-", delta_t_total, ".rds"))
# })

@@

vocc_plot <- plot_vocc(vocc,
  vec_aes = "distance",
  #max_vec_plotted = 100,
  min_vec_plotted = input_cell_size,
  NA_label = ".",
  fill_col = "diff",
  fill_label = paste("Change from\n", start_time, "to", start_time + delta_t_total, ""),
  raster_alpha = 1,
  white_zero = TRUE,
  vec_alpha = 0.25,
  axis_lables = FALSE,
  transform_col = no_trans#,
  #raster_limits = c(5, 13)
)
vocc_plot + ggtitle(climate)
}

@@

vocc_plot <- plot_vocc(vocc,
  vec_aes = "distance",
  #max_vec_plotted = 100,
  min_vec_plotted = input_cell_size,
  NA_label = ".",
  fill_col = "diff",
  fill_label = paste("Change from\n", start_time, "to", start_time + delta_t_total, ""),
  raster_alpha = 1,
  white_zero = TRUE,
  high_fill = "Steel Blue 4",
  low_fill = "Red 3",
  vec_alpha = 0.25,
  axis_lables = FALSE,
  transform_col = no_trans#,
  #raster_limits = c(5, 13)
)
vocc_plot + ggtitle(climate)
}

@@

vocc_plot_temp <- plot_vocc(vocc,
  vec_aes = "distance",
  #max_vec_plotted = 100,
  min_vec_plotted = input_cell_size,
  NA_label = ".",
  fill_col = "temp_diff",
  fill_label = paste("Change from\n", start_time, "to", start_time + delta_t_total, ""),
  raster_alpha = 1,
  white_zero = TRUE,
  vec_alpha = 0.25,
  axis_lables = FALSE,
  transform_col = no_trans#,
  #raster_limits = c(5, 13)
)+ ggtitle("temperature")

@@

vocc_plot_do <- plot_vocc(vocc,
  vec_aes = "distance",
  #max_vec_plotted = 100,
  min_vec_plotted = input_cell_size,
  NA_label = ".",
  fill_col = "do_diff",
  fill_label = paste("Change from\n", start_time, "to", start_time + delta_t_total, ""),
  raster_alpha = 1,
  white_zero = TRUE,
  high_fill = "Steel Blue 4",
  low_fill = "Red 3",
  vec_alpha = 0.25,
  axis_lables = FALSE,
  transform_col = no_trans#,
  #raster_limits = c(5, 13)
) + ggtitle("do")

@@

start <- biomass_change %>% 
  rename(start_density = est_exp_s, end_density = est_exp_e) %>% 
  select(icell, x, y, start_density, end_density)

end <- biomass_change %>% 
  rename(start_density = est_exp_s, 
         end_density = est_exp_e, 
         target_X = x, target_Y = y) %>% 
  select(icell, start_density, end_density, target_X, target_Y)

source_density <- biomass_change %>% 
  rename(source_density = est_exp_s) %>% 
  select(source_density, x, y)

# if (length(variable_names) == 1) {
#   
#   vectors <- vocc %>%
#     mutate(
#       start_cell = icell, 
#       vector_id = paste0(id, "_to_", tid),
#       initial_value = UQ(rlang::sym(paste0(variable_names, "_s"))), #var_1_s, 
#       source_end_value = UQ(rlang::sym(paste0(variable_names, "_e"))), #var_1_e, 
#       target_value = as.numeric(target_values)
#     ) %>%
#     filter(distance > 1)  %>%
#     select(vector_id, start_cell, x, y, target_X, target_Y, distance, 
#       initial_value, source_end_value, target_value, 
#       km_per_decade, timespan) 
#   
#   if (climate == "do") vectors <- vectors %>% 
#       filter(source_end_value < (optimal_do - do_threshold))
#   if (climate == "temperature") vectors <- vectors %>% 
#       filter(source_end_value > (optimal_temp + temp_threshold))
# 
# } else { # number of variables > 1
  # 
  # vectors <- vocc %>%
  #   mutate(
  #     start_cell = icell, vector_id = paste0(id, "_to_", tid),
  #     initial_value = s, source_end_value = e, target_value = target_values
  #   ) %>%
  #   filter(distance > 1) %>%
  #   select(vector_id, start_cell, x, y, target_X, target_Y, distance, 
  #     initial_value, source_end_value, target_value, 
  #     km_per_decade, timespan) 
  # 
  # # split source_end_value into multiple vectors
  # source_end <- purrr::map_df(strsplit(vectors$source_end_value, " "), 
  #                  function(x) data.frame(t(as.numeric(x))))
  # 
  # vectors <- cbind(vectors, source_end) 
  # 
  # for (j in seq_along(variable_names)){
  #   
  #   source_end_climatej <- paste0("X",j)
  #   vectors <- vectors %>% filter( 
  #     (!!sym(source_end_climatej)) < (optimal_value[j] + critical_thresholds[j])
  #     )
  # }
#  
#   #climate <- "do and temperature"
# # max_thresholds <- c(Inf, 0.75)
# # min_thresholds <- c(0.25, Inf)
# # critical_thresholds <- c(-0.25, 0.75)
# # optimal_value <- c(2, 6.5)
# #

  for (j in seq_along(variable_names)){

  vectors <- vocc %>%
    mutate(
      start_cell = icell,
      vector_id = paste0(id, "_to_", tid),
      source_end_value = e, # may need to turn into numeric vector later...
      target_value = target_values
    ) %>% filter(distance > 1)

      if (min_thresholds[j] != Inf) { 
        vectors <- filter(vectors, UQ(rlang::sym(paste0(variable_names[j], "_s"))) < optimal_value[j] - min_thresholds[j]) 
        }
      if (max_thresholds[j] != Inf) {
        vectors <- filter(vectors, UQ(rlang::sym(paste0(variable_names[j], "_s"))) > optimal_value[j] + max_thresholds[j])
        }

  #FIXME: decide which variables might cause problems in future merges and remove them...   
  #   vectors <- vectors %>% select(-N) 
  #

@@

# get mean and sd for densities without cell duplication
icells <- vocc_by_sp[[i]] %>% distinct(icell, .keep_all = TRUE)

@@

# calculate standardized density varibles
vocc_by_sp[[i]] <- vocc_by_sp[[i]] %>% 
  mutate(
    perc_change = (end_density - start_density)/start_density, 
    cell_type = as.factor(cell_type),
    #std_change = (perc_change)/sd.change, 
    std_source_density = (source_density - mean.density)/(sd.density*2), 
    mean_change = mean.change, sd_change = sd.change,
    mean_source_density = mean.density, sd_density = sd.density,
# set distance intercept at cell size
    distance = (ifelse(distance > 100, 100, distance)) - input_cell_size, 
    distance2 = distance*distance) %>% 
  rename(X = x, Y = y)

icells <- vocc_by_sp[[i]] %>% distinct(icell, .keep_all = TRUE)

mean.change <- mean(icells$perc_change, na.rm = TRUE)
sd.change <- sd(icells$perc_change, na.rm = TRUE)

vocc_by_sp[[i]] <- vocc_by_sp[[i]] %>% 
  mutate(std_change = (perc_change)/sd.change)

@@

vocc_by_sp[[i]] <- vocc_by_sp[[i]] %>% rename(
  source_end_climateA = X1, source_end_climateB = X2
  )

#browser()

icells <- vocc_by_sp[[i]] %>% distinct(icell, .keep_all = TRUE)
mean.climateA <- mean(icells$source_end_climateA, na.rm = TRUE)
sd.climateA <- sd(icells$source_end_climateA, na.rm = TRUE)
mean.climateB <- mean(icells$source_end_climateB, na.rm = TRUE)
sd.climateB <- sd(icells$source_end_climateB, na.rm = TRUE)

vocc_by_sp[[i]] <- vocc_by_sp[[i]] %>% 
  mutate(
    mean_climateA = mean.climateA, sd_climate = sd.climateA,
    std_end_climateA = (source_end_climateA - mean.climateA)/(sd.climateA*2),
    std_end_climateA2 = std_end_climateA^2,
    mean_climateB = mean.climateB, sd_climate = sd.climateB,
    std_end_climateB = (source_end_climateB - mean.climateB)/(sd.climateB*2),
    std_end_climateB2 = std_end_climateB^2
    )

#browser()
  end_climate <- purrr::map_df(strsplit(vocc_by_sp[[i]]$end_climate, " "), 
                   function(x) data.frame(t(as.numeric(x))))

  vocc_by_sp[[i]] <- cbind(vocc_by_sp[[i]], end_climate) %>%
    rename(end_climate_chr = end_climate, end_climate = X1, end_climateB = X2 )

 stop("Doesn't currently support testing vectors based on more than two climate variables at once.")
}
}

@@

# glimpse(vocc_by_sp[[i]])
# trim spatial outliers 
# Xmax <- quantile(vocc_by_sp[[i]]$X, 0.9995)
# Xmin <- quantile(vocc_by_sp[[i]]$X, 0.005)
# std_max <- quantile(vocc_by_sp[[i]]$std_change, 0.9995)
# 
# vocc_by_sp[[i]] <- vocc_by_sp[[i]] %>% filter ( X >= Xmin & X <= Xmax ) 
```

@@

Check distance threshold of 50 looks reasonable
```{r}
data <- vocc_by_sp[["adult"]]
#data <- filter(vocc_by_sp[["adult"]],  distance < 90) # distance > 1 &
#data <- filter(vocc_by_sp[["adult"]], distance > 1 ) #  &
